The inefficacy of superficial similarities for improving instructor-student relationships

### 2021-04-27 outputs/paper/paper.Rmd
why didn't it push the correctly-knitted version before

A scale is valid if it measures what it is supposed to measure [@Coolican_2014]. In this case, the scales should measure similarity perception and instructor-student relationship (ISR) perception. @citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers. Given the fact that these scales were developed for children, it is worth considering what the scales actually measure. To do this, we can look at the specific questions in the three distinct scales.

@@

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values. There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception). Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for intial perception and 3.5 versus 3.6 for end-of-term perception).

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale. Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced. Because this variable is measured differently than the two student similarity perception scales, and therefore modeled differently, direct comparisons between student and instructor perception of similarity should be made with caution.

@@

This study did successfully demonstrate relationship 1: the treatment caused a statistically significant increase in student similarity perception at the beginning and end of the term. Relationship 2 failed to materialize, as there was no significant difference in instructor similarity perception in the treatment group. To understand why the treatment effect was small for students and insignificant for instructors, we can consider the specifics of the treatment. What similarities were students and instructors informed of? [Appendix G] contains a complete list. Students and instructors were presented with generalities such as "An instructor’s most important characteristic? You both say being knowledgeable" or "You are cut from the same cloth–you both described yourself as thoughtful!" Such facts may not be unusual enough to be memorable, reducing the treatment's impact. It is also possible that similarities declared by an anonymous third party are simply not as impactful as similarities that are personally observed and experienced. Finally, because the instructor similarity perception was measured on one five-point scale, rather than a full scale as student similarity perception was, perhaps the measurement was less sensitive to change. @citeoriginal note that they used a single question to reduce the burden placed on instructors, who had to answer surveys for many students. However, this courtesy created a dissimilarity in measurment for outcomes that we would like to compare.

>>>>

A scale is valid if it measures what it is supposed to measure [@Coolican_2014]. In this case, the scales should measure similarity perception and instructor-student relationship (ISR) perception. @citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eighth-grade American students and their teachers. Given the fact that these scales were developed for children, it is worth considering what the scales actually measure. To do this, we can look at the specific questions in the three distinct scales.

@@

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values. There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception). Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for initial perception and 3.5 versus 3.6 for end-of-term perception).

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception. The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale. Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced. Because this variable is measured differently than the two student similarity perception scales, and therefore modeled differently, direct comparisons between student and instructor perception of similarity should be made with caution.

@@

This study did successfully demonstrate relationship 1: the treatment caused a statistically significant increase in student similarity perception at the beginning and end of the term. Relationship 2 failed to materialize, as there was no significant difference in instructor similarity perception in the treatment group. To understand why the treatment effect was small for students and insignificant for instructors, we can consider the specifics of the treatment. What similarities were students and instructors informed of? [Appendix G] contains a complete list. Students and instructors were presented with generalities such as "An instructor’s most important characteristic? You both say being knowledgeable" or "You are cut from the same cloth–you both described yourself as thoughtful!" Such facts may not be unusual enough to be memorable, reducing the treatment's impact. It is also possible that similarities declared by an anonymous third party are simply not as impactful as similarities that are personally observed and experienced. Finally, because the instructor similarity perception was measured on one five-point scale, rather than a full scale as student similarity perception was, perhaps the measurement was less sensitive to change. @citeoriginal note that they used a single question to reduce the burden placed on instructors, who had to answer surveys for many students. However, this courtesy created a dissimilarity in measurement for outcomes that we would like to compare.
